---
title: "CF_3"
output: pdf_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F, warning = F)
```

### Niko Hiananto

```{r}
setwd("C:/Users/nhian/Documents/UCLA MFE/Spring 2020/Computational Methods in Finance/Projects/R/Project 3")
##LGM
seed <- as.numeric(Sys.time())*100

lgm_unif <- function(n){
  m = 2^31 - 1
  a = 7^5
  
  #initialize time
  x0 = seed
  
  xn = numeric(n+1)
  xn[1] = x0
  #loop
  for(i in 2:length(xn)){
    xn[i] = (a * xn[i-1]) %% m
  }
  
  seed <<- tail(xn,1)
  
  unif = (xn+0.5) /m
  
  return(unif[-1])
}


##Box-Mueller
box_mueller <- function(n){
  out = n
  #if not even add 1 more since need 2 for each couple of normals
  n = ifelse(n %%2 == 0, n, n+1) 
  #generate n uniforms
  unif = lgm_unif(n)
  #create couples of normals for each couple of uniform
  z1 = sqrt(-2 * log(unif[1:(n/2)])) * cos(2 * pi *  unif[( (n/2)+1 ): length(unif)] )
  z2 = sqrt(-2 * log(unif[1:(n/2)])) * sin(2 * pi *  unif[( (n/2)+1 ): length(unif)] )
  #return only n number of normals
  return( c(z1,z2)[1:out])
}

#function to find the nth root
nroot <- function(x, n){
    sign(x) * abs(x)^(1/n)
}
```

### Question 1. Expected Values and Probabilities
```{r}
#1.

x0 = 1
y0 = 3/4

#create function for dx
dx <- function(x, dt, dz){
  (0.2 - 0.5*x)*dt + (2/3) * sqrt(dt) * dz
}

#create function for dy
dy <- function(y, dt, dz, t){
  (2/(1+t)*y + (1 + t^3)/3)*dt + ((1 + t^3)/3 * sqrt(dt) * dz)
}

fx_q1 <- function(x0, y0, paths = 1000, dt = 0.01){
  #X0-X2  
  ncolx = 2/dt
  ncoly = 3/dt
  step = 1/dt
  
  
  x = matrix(0, nrow = paths, ncol = ncolx + 1) 
  #Y0-Y3
  y = matrix(0, nrow = paths, ncol = ncoly + 1) 
  x[,1] = x0
  y[,1] = y0
  
  for(i in 2:ncol(x)){
    dz = box_mueller(paths)
    x[,i] = x[,i-1] + dx(x[,i-1], dt = dt, dz = dz)
  }
  
  for(i in 2:ncol(y)){
    dz = box_mueller(paths)
    y[,i] = y[,i-1] + dy(y[,i-1], dt = dt, dz = dz, t = (i-2) * dt )
  }
  
  
  p1 = sum(y[, 2 * step + 1] > 5)/nrow(y)
  e1 = mean(nroot((x[,2 * step + 1]),3))
  e2 = mean(y[,3 * step + 1])
  e3 = mean(x[,2 * step + 1] * y[,2 *step + 1] * (x[,2 * step + 1] > 1))
  
  return(c("p1" = p1,
           "e1" = e1,
           "e2" = e2,
           "e3" = e3))
}


q1_res = fx_q1(x0,y0, 10000, 0.01);q1_res
```
The expected values and probabilities with 10,000 paths and 0.01 step size is: \
```{r}
q1_res
```

\

### Question 2. Expected Values
```{r}
#2.

#create function for dx
dx_2 <- function(x, dt, dw, dz){
  0.25 * x * dt + ( (1/3)*x* sqrt(dt) * dw ) - (0.75 * x * sqrt(dt) * dz)
}

#create function for y
y_2 <- function(t, dw, dz){
  exp(-0.08 * t + (1/3) * sqrt(t) * dw + 0.75 * sqrt(t) * dz )
}


fx_q2 <- function(x0, paths = 1000, dt = 0.01){
  n = 3/dt
  steps = 1/dt
  
  #X0-X3
  x = matrix(0, nrow = paths, ncol = n + 1)
  x[,1] = x0
  
  #Y1
  dz = matrix(box_mueller(2 * paths), ncol = 2)
  y = y_2(1, dz[,1], dz[,2])
  
  
  for(i in 2:ncol(x)){
    dz = matrix(box_mueller(2*paths), ncol = 2)
    x[,i] = x[,i-1] + dx_2(x[,i-1], dt = dt, dw = dz[,1], dz = dz[,2])
  }
  
  e1 = mean(nroot((1 + x[, 3 * steps + 1]),3))
  e2 = mean(x[,1 * steps + 1] * y)

  return(c("e1" = e1,
           "e2" = e2))
  
}

q2_res <- fx_q2(1, 10000, 0.01) 
```
The resulting expected values for e1 and e2 with 10,000 paths and 0.01 step size is: \
```{r}
q2_res
```
\

### Question 3 Monte Carlo Option Pricing
```{r}
#3 MC Call Option
c_price_mc_var <- function(s0, X, t, r, sigma, paths = 10000){
  rand_norm = box_mueller(ceiling(paths/2))
  
  s_t_p = s0 * exp(sigma * rand_norm * sqrt(t) + (r - (sigma^2)/2) * t)
  s_t_m = s0 * exp(sigma * -rand_norm * sqrt(t) + (r - (sigma^2)/2) * t) #anti-thetic variate
  
  payoff_p = pmax(s_t_p - X, 0)
  payoff_m = pmax(s_t_m - X, 0)
  payoff = (payoff_p + payoff_m)/2
  
  prc = exp(-r * t) * mean(payoff)
  return(prc)
}


sim_prc_var <- c_price_mc_var(20, 20, 0.5, 0.04, 0.25)
```
(a.) The European option price with parameters $S_0 = 20$, $X = 20$, $\sigma = 0.25$, $r = 0.04$, $T = 0.5$ and antithetic variate reduction technique with 10,000 paths is shown below: \
```{r}
sim_prc_var
```

```{r}
#b. BS formula
approx_n <- function(x){
  d = c(0.0498673470,
        0.0211410061,
        0.0032776263,
        0.0000380036,
        0.0000488906,
        0.0000053830)
  
  n_x <- function(x){
    1 - 0.5 * ((1 + sum(d * x^(1:6)) )^(-16))
  }
  
  res = ifelse(x < 0, 1 - n_x(-x), n_x(x))
  return(res)
}


bs_formula <- function(s, X, t, r, sigma){
  d1 = 1/(sigma * sqrt(t)) * (log(s/X) + ((r + (sigma^2) / 2 )* t))
  d2 = d1 - (sigma * sqrt(t))
  c = s * approx_n(d1) - X * approx_n(d2) * exp(- r * t)
  return(c("prc"=c))
}

bs_prc = bs_formula(20, 20, 0.5, 0.04, 0.25)
```
(b.) The price of the call option using the Black-Scholes formula and with the same parameters ($S_0 = 20$, $X = 20$, $\sigma = 0.25$, $r = 0.04$, $T = 0.5$) and the approimation of $N(.)$ is shown below: \
```{r}
bs_prc
```

```{r}
#c. estimating greeks
est_greeks <- function(s0, k, t, r, sigma){
  #delta
  
  delta = (bs_formula(s0+0.01, k, t, r, sigma) - bs_formula(s0-0.01, k, t, r, sigma))/(2*0.01)
  
  #gamma
  gamma = (bs_formula(s0+ (2 * 0.01), k, t, r, sigma) - 2*bs_formula(s0, k, t, r, sigma) + bs_formula(s0- (2 * 0.01), k, t, r, sigma))/(4 * 0.01^2)
  
  #theta
  theta = (bs_formula(s0, k, t-0.004, r, sigma) - bs_formula(s0, k, t+0.004, r, sigma))/(2 * 0.004)
  
  #vega
  vega = (bs_formula(s0, k, t, r, sigma+0.001) - bs_formula(s0, k, t, r, sigma-0.001))/(2*0.001)
  
  return(c("delta" = delta,
           "gamma" = gamma,
           "theta" = theta,
           "vega" = vega))
}

s_list = 15:25
greeks = simplify2array(lapply(s_list, function(x) est_greeks(x, 20 , 0.5, 0.04,0.25)))
```
(c.) The plot for the greeks with stock prices 15:25 is shown by the figures below:
```{r, fig.align = "center", fig.width = 6, fig.height=4}
{par(mfrow = c(1,1))
plot(x = s_list, y = greeks[1,], main = "Delta", xlab = "Spot Price", ylab = "Delta", type = 'l')
plot(x = s_list, y = greeks[2,], main = "Gamma", xlab = "Spot Price", ylab = "Gamma", type = 'l')
plot(x = s_list, y = greeks[3,], main = "Theta", xlab = "Spot Price", ylab = "Theta", type = 'l')
plot(x = s_list, y = greeks[4,], main = "Vega", xlab = "Spot Price", ylab = "Vega", type = 'l')}
```


### Question 4. Heston Model
```{r}
#4. Heston Model
heston <- function(s0, K, t, r, sigma, a, b, v0 = 0.05, rho, paths = 10000, dt = 1){
  #years
  n = t/dt 
  
  sim <- function(paths, fun, ...){
    #fun = function to pass thru the different methods
    #create initial matrix data to store results/simulations
    s = v = matrix(0, ncol = n+1, nrow = paths)
    #initial values
    v[,1] = v0
    s[,1] = s0
    for(i in 2:ncol(v)){
      #create correlated normals
      dz = matrix(box_mueller(paths * 2), ncol = 2)
      dz[,2] = rho * dz[,1] + sqrt(1 - (rho^2)) * dz[,2]
      
      #simulate volatility path
      v[,i] = fun(v[,i-1], 1) + (a * (b - fun(v[i-1], 2) ) * dt ) + 
        (sigma * sqrt( fun(v[,i-1],3) ) * sqrt(dt) * dz[,2])
      
      #simulate stock path
      s[,i] = s[,i-1] + (r * s[,i-1] * dt) + (sqrt( fun(v[,i-1], 3) ) * s[,i-1] * sqrt(dt) * dz[,1])
      
    }
    
    return( exp(-r *t) * mean( pmax( (s[,ncol(s)] - K), 0), na.rm = T ) )
  }
  
  #create the 3 methods
  reflection <- function(x, sch = 1) abs(x)
  partial <- function(x, sch = 1){
    if(sch == 1){
      return(x)
    } else if (sch == 2){
      return(x)
    } else if (sch == 3){
      return(pmax(0,x))
    }
  }
  full <- function(x, sch = 1){
    if(sch == 1){
      return(x)
    } else if (sch == 2){
      return(pmax(0,x))
    } else if (sch == 3){
      return(pmax(0,x))
    }
  }
    
  #pass through the methods
  c1 = sim(paths, reflection)
  c2 = sim(paths, partial)
  c3 = sim(paths, full)
  return(c("c1" = c1,
           "c2" = c2,
           "c3" = c3))
}

s = 48
k = 50
t = 1
r = 0.03
v0 = 0.05
sigma = 0.42
a = 5.8
beta = 0.0625
rho = -0.6

res_heston <- heston(s0 = s, K = k, t = t, r = r, sigma = sigma, a = a,
       b = beta, v0 = v0, rho = rho, paths = 100000, dt = 0.01)

```
The price of a European Call using the Heston Model wiith 100,000 paths and step size = 0.01 for the reflection (c1), partial truncation (c2) and full truncation (c3) method respectively is shown below: \
```{r}
res_heston
```


### Question 5. Quasi Monte-Carlo

```{r}
#5 Quasi-MC
#a.
unif_2d = matrix(lgm_unif(200), ncol = 2)

#b. halton
halton_seq = function(n, base) {
  res_vec = numeric(n)
  
  for(index in 1:n){
    fraction = 1
    result = 0
    ct = index
    while (index > 0) {
      fraction = fraction / base
      result = result + fraction * (index %% base)
      index = floor(index / base)
    }
    res_vec[ct] = result
  }
  return(res_vec)
}
#2d halton
halton_2d_1 <- simplify2array(lapply(c(2,7), function(x) halton_seq(n = 100, base = x)))

#common factor
halton_2d_2 <- simplify2array(lapply(c(2,4), function(x) halton_seq(n = 100, base = x)))
```
(d.) The plots for all three 2-D generated sequences is shown below: \
```{r, fig.align = "center"}
#plots
{par(mfrow = c(3,1))
plot(x = unif_2d[,1], y = unif_2d[,2], xlab = "x", ylab = "y", main = "Uniform")
plot(x = halton_2d_1[,1], y = halton_2d_1[,2], xlab = "x", ylab = "y", main = "Halton (2,7)")
plot(x = halton_2d_2[,1], y = halton_2d_2[,2], xlab = "x", ylab = "y", main = "Halton (2,4)")
}
```
The Halton(2,7) with the prime number bases show the most uniformly distributed sequence which makes it the most ideal to use to approximate the value of integrals. The sequences generated from the uniform distribution are quite spread out nicely but the Haston sequence are still ahead in terms of evenly spread out points. The Halton (2,4), however, has the same common factor and thus suffers from it which results in a non-uniformly spread out points and instead forms a certain region where the sequences are most concentrated at. \
\
(e.) Using the 2-D Halton and 10,000 number of sequences, the estimates of the integral for the different bases (2,7), (2,4), (5,7) are given by: \
```{r}
#integral estimation
int_fx <- function(x,y){
  (exp(-x * y) * (sin(6 * pi *x) + nroot(cos(2 * pi * y),3) ) )
}

seq_num = 10000
estimate_int <- function(bases){
  #bases a vector of 2
  int_seq = simplify2array(lapply(bases, function(x) halton_seq(n = seq_num, base = x)))
  res = sum(mapply(function(x, y) int_fx(x, y), int_seq[,1], int_seq[,2]))/(seq_num)
  return(res)
}

#integral calculation
res = lapply(list("2,7" = c(2,7), "2,4" = c(2,4), "5,7" = c(5,7)), function(x) estimate_int(x))
res
```

